--- 
title: "DTS Processing"
author: "G360"
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
documentclass: krantz
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
fontsize: 12pt
monofont: "Source Code Pro"
monofontoptions: "Scale=0.7"
site: bookdown::bookdown_site
description: "This report provides information about an A-DTS test."
params:
  input: NA
  dir_name: NA
  heating_power: 15
  heating_type: 'heating'
  has_coldspray: TRUE
  n_groups: 3
  zero_distance: 0
---

```{r setup, include=FALSE}
print(getwd())
knitr::opts_chunk$set(echo = FALSE, comment = NA)

library(data.table)
library(ggplot2)
library(dts)
library(plotly)
library(viridis)
library(irlba)
library(mclust)
library(knitr)
library(gganimate)
library(transformr)
library(bookdown)
library(DT)
library(htmlwidgets)
library(knitr)
library(widgetframe)
library(svglite)
library(gifski)
library(aquifer)

a <- Sys.time()

dts           <- params$input
heating_type  <- params$heating_type
heating_power <- params$heating_power
n_groups      <- params$n_groups
has_coldspray <- params$has_coldspray
zero_distance <- params$zero_distance


```

# Introduction

This script generates a summary report for A-DTS for a single borehole. The only input necessary is the path to the folder that contains XML files. This is a test change.


File path: `r params$dir_name`

```{r getdata, warning=FALSE, message = FALSE, include = FALSE}

distance_ranges <- function(x, by = 'bath') {
  get_distance_table(x)[, list(min_distance = min(distance),
                      max_distance = max(distance)), by = list(type = get(by))]
}
time_ranges <- function(x, by = 'type') {
  get_time_table(x)[, list(min_time = min(start),
                      max_time = max(start)), by = list(type = get(by))]
}

# smooth for calibration - not sure this helps

# attempt to reduce noise
dts$trace_time$calib_temperature <- caTools::runmean(dts$trace_time$calib_temperature, 6, endrule = 'keep')

dts <- denoise_by_trace(dts)


# correlation to probe temperature method to locate the water bath
dts <- find_water_bath(dts)

# apply the calibration to data
dts <- bath_calibration(dts)


# limit the distance range to aid in finding heating cutoff
distance_range <- get_bath_limits(dts)
dts <- heating_time(dts, 
                    heating_type = heating_type, 
                    distance_range = distance_range)

dts <- heating_distance(dts, 
                        heating_type = heating_type)


# half <- combine_by_symmetry(dts)
half <- half_data(dts, type='head')
  
half <- find_coldspray(half, type = 'head')

if (has_coldspray) {
  half <- set_zero_distance(half)
} else {
  half <- set_zero_distance(half, zero_distance)
}



half <- add_heating_cooling(half)

# the below functions could be improved
classes <- heated_groups(subset_distance(half, by = 'borehole'), 
                      n_comp = 5, 
                      n_clust = n_groups,
                      n_buffer = 1000)

classes$trace_distance[, classification := process_classification(classification, k = 31)]
classes$trace_distance[, classification := process_terminal(classification, k = 15)]
classes    <- add_classification(classes)


half_60 <- copy(half) #average_time(half, n = 120)


# for report
coldspray_location <- half$trace_distance[borehole == TRUE]$distance[1]
n_files  <- nrow(dts$trace_time)
time     <- range(get_times(dts))
distance <- range(unique(dts$trace_distance$distance))
delta_distance <- mean(diff(half$trace_distance$distance), na.rm=TRUE)
delta_time <- mean(diff(half$trace_time$mid), na.rm=TRUE)


# decrease memory consumption
rm(dts) 
gc()

```


- Total number of files processed: `r n_files`
- Time Range: `r time[1]` to `r time[2]`
- Mean sample interval (time) `r round(delta_time, 2)`
- Distance Range: `r distance[1]` to `r distance[2]`
- Mean sample interval (depth) `r round(delta_distance, 2)`
- Borehole starts: `r coldspray_location`


```{r oupputdata, eval = FALSE}


write_dts(dts)

# write matrix
# write apparent thermal conductivity
# write averages


```



# Reference temperatures

## Raw


```{r reftemperature, warning=FALSE}

tr_time <- time_ranges(half, by = 'type')
tr_time[, ymin := -Inf]
tr_time[, ymax :=  Inf]

meta <- half$trace_time
meta <- melt(meta[, list(start, probe_1, probe_2, ref_temperature, bath_temp_dts)],
             id.vars = 'start')
meta[, value_adj := value - frollmean(value, 10), by = variable]

p <- ggplot() +
  geom_rect(data = tr_time, aes(xmin = min_time,
                                xmax = max_time,
                                ymin = ymin,
                                ymax = ymax,
                                fill  = type),
            color = 'transparent') +
  scale_fill_manual(values = c('#fbc93d10', '#6cc0e510', '#fb4f4f10'),
                    breaks = c('ambient', 'cooling', 'heating')) +
  geom_line(data = meta, aes(x = start, y = value)) +
  scale_x_datetime(expand = c(0,0)) +
  facet_wrap(variable~., ncol = 1, scales = 'free_y') +
  ylab('Temperature (°C)') +
  theme_bw() +
  theme(legend.position = 'top',
        axis.title.x = element_blank())
p

```


## Detrended

```{r reftemperaturedemeaned, warning=FALSE}

p <- ggplot() +
  geom_rect(data = tr_time, aes(xmin = min_time,
                           xmax = max_time,
                           ymin = ymin,
                           ymax = ymax,
                           fill  = type),
            color = 'transparent') +
  scale_fill_manual(values = c('#fbc93d10', '#6cc0e510', '#fb4f4f10'),
                     breaks = c('ambient', 'cooling', 'heating')) +
  geom_line(data = meta, aes(x = start, y = value_adj)) +
  scale_x_datetime(expand = c(0,0))+
  facet_wrap(variable~., ncol = 1) +
  ylab('Temperature (°C)') +
  theme_bw() +
  theme(legend.position = 'top',
        axis.title.x = element_blank())
p



```


## Adjustment


```{r reftemperatureadjust, warning=FALSE}

meta <- get_time_table(half)
meta <- melt(meta[, list(start,
                         probe_mean = calib_temperature,
                         dts_bath_mean = bath_temp_dts,
                         calib_adj = -calib_adj,
                         dts_adjusted = bath_temp_dts - calib_adj)],
             id.vars = 'start')

mean_t <- meta[, list(t_mean = paste('mean: ', round(mean(value, na.rm = TRUE),2), '(°C)'),
                      start = min(start)), by = variable]



meta[, value_adj := value-mean(value, na.rm = TRUE), by = variable]
p <- ggplot() +
  geom_rect(data = tr_time, aes(xmin = min_time,
                           xmax = max_time,
                           ymin = ymin,
                           ymax = ymax,
                           fill  = type),
            color = 'transparent') +
  geom_label(data = mean_t, aes(x = start, y = Inf, label = t_mean),
                    color = viridis(1), size = 4, vjust = "inward", hjust = "inward",
                    label.padding = unit(0.1, units = 'lines')) +
  scale_fill_manual(values = c('#fbc93d10', '#6cc0e510', '#fb4f4f10'),
                     breaks = c('ambient', 'cooling', 'heating')) +
  geom_line(data = meta, aes(x = start, y = value_adj)) +
  scale_x_datetime(expand = c(0, 0))+
  facet_wrap(variable~., ncol = 1) +
  ylab('Temperature (°C)') +
  theme_bw() +
  theme(legend.position = 'top',
        axis.title.x = element_blank())
p
```


**Smoothing the probe signal prior to adjustment may slightly decrease noise as the probes have more noise than expected.**

dts_adjusted is the same as the probe_mean with this method.


Water bath mean:

$\overline{T}_{t}^{bath} = \frac{\Sigma T_{t}^{bath}}{n}$

Probe mean:

$\overline{T}_{t}^{probes} = \frac{T_{t}^{probe 1} + T_{t}^{probe 2}}{2}$

Adjustment:

$\overline{T}_{t}^{adj} =  \overline{T}_{t}^{probes} - \overline{T}_{t}^{bath}$


# Key locations and times (autopicks)

## Water bath

```{r waterbathtable}

sub_trace <- sample_times(half, n_traces = 3)
tr <- distance_ranges(half, by = 'bath')
tr[, ymin := -Inf]
tr[, ymax :=  Inf]

```

Distances `r tr[type== TRUE]$min_distance` to `r tr[type== TRUE]$max_distance` are likely in a water bath.

```{r waterbath}




p <- ggplot() +
  geom_rect(data = tr, aes(xmin = min_distance,
                           xmax = max_distance,
                           ymin = ymin,
                           ymax = ymax,
                           fill  = type),
            color = 'transparent') +
  geom_line(data = sub_trace, aes(x = distance, y = temperature)) +
  coord_flip() +
  scale_x_reverse(expand = c(0,0)) +
  scale_fill_manual(values = c('#6cc0e510', '#fb4f4f10'),
                     breaks = c(FALSE, TRUE)) +
  facet_wrap(as.character(start)~., ncol = 5) +
  theme_bw() +
  xlab('Distance (m)') +
  ylab('Temperature (°C)') +
  theme(legend.position = 'top',
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        axis.text = element_text(size = 10),
        strip.text = element_text(size = 10,
                                  margin = margin(0,0,0,0, "cm")))

p


```

```{r waterbathmatrix}

rng <- range(half_60$trace_distance[bath == TRUE]$distance)
sub_60 <- subset_distance(half_60, rng[1], rng[2])
m <- to_matrix(sub_60)

p <- plot_ly(z = m,
             x = as.POSIXct(as.numeric(colnames(m)), origin = '1970-01-01', tz = 'UTC'),
             y = rownames(m),
             colors = viridis(100),
             type = "heatmap",
             height = 200) |>
    layout(yaxis = list(autorange="reversed"))
p

```

## Ambient Heating Cooling Times

```{r ambientheatingcoolingtable}
kable(tr_time[, list(type, min_time, max_time)])
```

```{r ambientheatingcooling, dev = 'png'}

sub_trace <- sample_distance(half, n_traces = 20)

p <- ggplot() +
  geom_rect(data = tr_time, aes(xmin = min_time,
                           xmax = max_time,
                           ymin = ymin,
                           ymax = ymax,
                           fill  = type),
            color = 'transparent') +
  geom_line(data = sub_trace, aes(x = start, y = temperature)) +
  scale_x_datetime(expand = c(0,0)) +
  scale_fill_manual(values = c('#fbc93d10', '#6cc0e510', '#fb4f4f10'),
                     breaks = c('ambient', 'cooling', 'heating')) +
  facet_wrap(distance~., scales = 'free_y', ncol = 5) +
  theme_bw() +
  xlab('') +
  ylab('Temperature (°C)') +
  theme(legend.position = 'top',
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        axis.text = element_text(size = 6),
        axis.title.x = element_blank(),
        strip.text = element_text(size = 4,
                                  margin = margin(0,0,0,0, "cm")))

p

```


## Ambient Heating Distances

```{r ambientheatingcoolingdistancestable}

tr <- distance_ranges(half, by = 'heated')
tr[, ymin := -Inf]
tr[, ymax :=  Inf]

sub_trace <- sample_times(half, n_traces = 10)

kable(tr[, list(type = ifelse(type, 'Heated', 'Not heated'), min_distance, max_distance)])

```


```{r ambientheatingcoolingdistances}

p <- ggplot() +
  geom_rect(data = tr, aes(xmin = min_distance,
                           xmax = max_distance,
                           ymin = ymin,
                           ymax = ymax,
                           fill  = type),
            color = 'transparent') +
  geom_line(data = sub_trace, aes(x = distance, y = temperature)) +
  coord_flip() +
  scale_x_reverse(expand = c(0,0)) +
  scale_fill_manual(values = c('#6cc0e510', '#fb4f4f10'),
                     breaks = c(FALSE, TRUE)) +
  facet_wrap(as.character(start)~., ncol = 5) +
  theme_bw() +
  xlab('Distance (m)') +
  ylab('Temperature (°C)') +
  theme(legend.position = 'top',
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        axis.text = element_text(size = 10),
        strip.text = element_text(size = 6,
                                  margin = margin(0,0,0,0, "cm")))

p

```



# Classification

## Figure

```{r other}
classes_60 <- copy(classes) #average_time(classes, n = 120)


sub_trace <- sample_times(classes_60, n_traces = 5)

p <- ggplot() +
  geom_point(data = na.omit(sub_trace),
             aes(x = distance,
                 y = temperature,
                 color = as.factor(classification)),
             size = 0.8) +
  coord_flip() +
  scale_x_reverse(expand = c(0,0)) +
  facet_wrap(as.character(start)~., ncol = 5) +
  scale_color_brewer(type = 'qual', palette = 'Dark2', name = 'Class') +
  theme_bw() +
  xlab('Distance (m)') +
  ylab('Temperature (°C)') +
  theme(legend.position = 'top',
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(0.1, "lines"),
        axis.text = element_text(size = 10),
        strip.text = element_text(size = 10,
                                  margin = margin(0,0,0,0, "cm")))

p


```

## Table

```{r classificationmatrix}

rng <- range(classes_60$trace_distance[heated == TRUE]$distance)
sub_60 <- subset_distance(classes_60, rng[1], rng[2])

subs <- na.omit(unique(sub_60$trace_distance$classification))

am <- range(classes_60$trace_time[type=='ambient']$start)
if(heating_type %in% c('heating', 'both')){
  he <- range(classes_60$trace_time[type=='heating']$start)
}
if(heating_type %in% c('cooling', 'both')){
  co <- range(classes_60$trace_time[type=='cooling']$start)
}

plot_group <- function(x, group = n_groups, n_trim = 1) {
  rng <- range(x$trace_distance[classification %in% group]$distance)
  d <- subset_distance(x, rng[1], rng[2])
  m <- to_matrix(d)
  
  if(ncol(m) < n_trim * 3) {
    n_trim <- 1
  }
  
  m <- m[, -c(1:n_trim, ncol(m):(ncol(m)-n_trim+1))]
  x_lab <- as.POSIXct(as.numeric(colnames(m)),
                      origin = '1970-01-01', tz = 'UTC')
  y_lab <- rownames(m)
  plot_ly(z = (m),
          x = x_lab,
          y = y_lab,
          colors = viridis(100),
          type = "heatmap") |>
    layout(yaxis = list(autorange="reversed"))
}


```


```{r classificationmatrixtable}

d <- classes_60$trace_distance
d <- d[, list(min_distance = min(distance, na.rm = TRUE),
         max_distance = max(distance, na.rm = TRUE)), by = classification]

kable(d[classification %in% c(1:n_groups, NA)])

```

# Images

Would log colorscales be more useful than linear ones when large changes are expected?

## All time



```{r matall}
p <- list()
for (i in seq_along(subs)) {
  p[[i]] <- plot_group(sub_60, subs[i]) |>
  colorbar(title = "Temperature", x = 1, y = (1 - (i-1) / length(subs)))
}

subplot(p, nrows = length(subs), shareX = TRUE)


```

## Ambient

```{r am}
p <- list()
for (i in seq_along(subs)) {
  p[[i]] <- plot_group(subset_time(sub_60, am[1], am[2]), subs[i]) |>
  colorbar(title = "Temperature", x = 1, y = 1-(i-1)/length(subs))
}

subplot(p, nrows = length(subs), shareX = TRUE)

```

## Heating


```{r he}
if(heating_type %in% c('heating', 'both')){

p <- list()
for (i in seq_along(subs)) {
  p[[i]] <- plot_group(subset_time(sub_60, he[1], he[2]), subs[i]) |>
  colorbar(title = "Temperature", x = 1, y = 1-(i-1)/length(subs))
}

subplot(p, nrows = length(subs), shareX = TRUE)
} else {
  print('no heating')
}
```


## Cooling

```{r co}
if(heating_type %in% c('cooling', 'both')){
  
p <- list()
for (i in seq_along(subs)) {
  p[[i]] <- plot_group(subset_time(sub_60, co[1], co[2]), subs[i]) |>
  colorbar(title = "Temperature", x = 1, y = 1-(i-1)/length(subs))
}

subplot(p, nrows = length(subs),, shareX = TRUE)
} else {
  print('no cooling')
}

```


## Saturated ambient

```{r amtmp}
plot_group(subset_time(sub_60, am[1], am[2]), group = n_groups) |>
  colorbar(title = "Temperature")

```

## Saturated heating

```{r hetmp}
if(heating_type %in% c('heating', 'both')){
  plot_group(subset_time(sub_60, he[1], he[2]), group = n_groups) |>
    colorbar(title = "Temperature")
} else {
  print('no heating')
  
}
```

## Saturated cooling

```{r cotmp}
if(heating_type %in% c('cooling', 'both')){
  plot_group(subset_time(half_60, co[1], co[2]), group = n_groups) |>
    colorbar(title = "Temperature")
} else {
  print('no cooling')
  
}

```

# Slope (Linear fit to log-linear data)

A linear model is fit to the heating and cooling data.  The two results are compared.

## Trim late times

The length of the series goes from 30 seconds to the facet value.

```{r }

max_heat_time <- max(classes$trace_time[type=='heating']$elapsed_time, na.rm = TRUE)

et <- c(60, 300, 900, 1800, 3600)
et <- et[et < max_heat_time]

ff_list <- list()
fb_list <- list()
for (i in seq_along(et)) {
  ff <- fit_heating_cooling(classes, start_time = 30, end_time = et[i], heating_type = heating_type)
  rng <- range(classes$trace_distance[classification %in% n_groups]$distance)
  ff <- subset_distance(ff, rng[1], rng[2])
  ff[, start_time := 30]
  ff[, end_time := et[i]]
  ff[, method := 'forward']
  ff_list[[i]] <- copy(ff)
}

et <- 1:5 * 7200
et <- et[et < max_heat_time]

for (i in seq_along(et)) {
  start_time <- ((max_heat_time-et[i]) %/% 1800) * 1800
  fb <- fit_heating_cooling(classes, start_time = start_time, end_time = max_heat_time, heating_type = heating_type)
  rng <- range(classes$trace_distance[classification %in% n_groups]$distance)
  fb <- subset_distance(fb, rng[1], rng[2])
  fb[, start_time :=  start_time]
  fb[, end_time := max_heat_time]
  fb[, method := 'backward']
  fb_list[[i]] <- copy(fb)

}

fits <- rbindlist(list(rbindlist(fb_list), rbindlist(ff_list)))

fits[, thermal_conductivity := thermal_conductivity(abs(slope), heating_power)]
p <- ggplot(fits[method == 'forward'], aes(y = abs(slope), x = distance, color = type)) +
  geom_line() +
  scale_x_reverse() +
  scale_color_manual(values = c('#4682B4', '#b44682')) +
  coord_flip() +
  facet_wrap(end_time~., ncol = 5) +
  theme_bw()

p

```


## Trim early times

The length of the series goes from the facet value to `r format(max_heat_time, scientific = FALSE)` seconds.


```{r }

p <- ggplot(fits[method == 'backward'], aes(y = abs(slope), x = distance, color = type)) +
  geom_line() +
  scale_x_reverse() +
  scale_color_manual(values = c('#4682B4', '#b44682')) +
  coord_flip() +
  facet_wrap(start_time~., ncol = 5) +
  theme_bw()

p

```



<!-- # Thermal Conductivity (distributed lag deconvolution) -->

<!-- By using deconvolution in time the effective test length of the test can be extended. -->

<!-- Question: What causes the diminished apparent thermal conductivity with time? -->

<!-- - Transfer from above/below in formation? -->

<!-- - Heat transfer in liner water? -->

<!-- - How can one choose the appropriate slope for quantitative results? -->

<!-- Question: How are the appropriate rates selected to estimate properties of different components of the system? -->

<!-- - Flow -->

<!-- - Vertical thermal conductivity -->

<!-- - Bulk thermal conductivity -->

<!-- - Porosity estimation -->


<!-- ## Thermal conductivity at elapsed times -->

<!-- ```{r message = FALSE, warning = FALSE} -->
<!-- # n_sub <- 5 -->
<!-- # half_05 <- average_time(half, n_sub) -->
<!-- # fits <- fit_convolve(half_05, n_knots = 6, cool_mult = 2) -->
<!-- #  -->
<!-- # y <- seq(n_sub, nrow(fits)*n_sub, by = n_sub) -->
<!-- #  -->
<!-- # log_deriv <- function(x, y) { -->
<!-- #   x / (log(data.table::shift(y, n = 1, type = 'lead')) - log(data.table::shift(y, n = 1, type = 'lag'))) -->
<!-- # } -->
<!-- # ld <- apply(fits, 2, function(x) log_deriv(x, y)) -->
<!-- # ld[is.na(ld)] <- 0 -->
<!-- # fits <- apply(ld, 2, function(x) aquifer::fftw_convolve(x, waterlevel::window_rectangular(301))) -->
<!-- # fits <- cbind(get_distance_table(half_05)[, list(distance, classification)], t(fits)) -->
<!-- #  -->
<!-- # fits <- fits[classification == 6] -->
<!-- # fits[, classification := NULL] -->
<!-- #  -->
<!-- # fits <- melt(fits, id.vars = 'distance') -->
<!-- # fits[, variable := as.numeric(gsub('V', '', variable)) * 5] -->
<!-- # fits[, thermal_conductivity := thermal_conductivity(value, 7.5)] -->
<!-- #  -->
<!-- # p <- ggplot(fits[variable %in% seq(1800, 3600*8, 300)], -->
<!-- #             aes(y = thermal_conductivity, x = distance)) + -->
<!-- #   geom_line() + -->
<!-- #   geom_hline(yintercept = 4.6, color = 'red') + -->
<!-- #   scale_x_reverse() + -->
<!-- #   scale_y_continuous(limits = c(0, 100)) + -->
<!-- #   scale_color_viridis_c() + -->
<!-- #   coord_flip() + -->
<!-- #   theme_bw() + -->
<!-- #   transition_states(variable) + -->
<!-- #   ease_aes('cubic-in-out') +  -->
<!-- #   labs(title = "Slope at elapsed time: {round(frame * 300 + 1500, 0)}") -->
<!-- #  -->
<!-- # options(gganimate.device = 'png') -->
<!-- # options(gganimate.nframes = 91) -->
<!-- # options(gganimate.fps = 10) -->
<!-- # options(gganimate.renderer = ffmpeg_renderer()) -->
<!-- # options(gganimate.dev_args = list(width = 600, height= 500)) -->
<!-- # animate(p, device = 'png') -->

<!-- ``` -->


<!-- ```{r, dev = 'png'} -->
<!-- #  -->
<!-- # p <- ggplot(fits[variable %in% seq(1800, 3600*8, 20)], aes(y = thermal_conductivity, x = variable, color = distance, group = distance)) + -->
<!-- #   geom_line() + -->
<!-- #   scale_color_viridis_c(alpha = 0.5) + -->
<!-- #   annotation_logticks(sides = 'bt')+ -->
<!-- #   scale_x_log10(expand = c(0,0)) + -->
<!-- #   xlab('Elapsed Time (Seconds)') + -->
<!-- #   ylab('Thermal Conductivity (W/mK)') + -->
<!-- #   theme_bw() -->
<!-- # p -->
<!-- ``` -->

# Footer

Creation Date: `r Sys.time()`

```{r}

sessionInfo()

```


```{r}
b <- Sys.time()

cat(paste('Processing time: ', round(as.numeric(b)-as.numeric(a), 0), 'seconds'))

```



<!--chapter:end:index.Rmd-->

